---
title: "alndist"
author: "Yue Jiang"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{alndist}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
library(alndist)
```

We use aligned mouse olfactory receptors as example. There are about 1000 of them. Aligned fasta file is provided as extdata in this package. Load these sequences.

```{r}
or <- seqinr::read.fasta(system.file("extdata/mouseOR.fasta", package="alndist"), seqtype="AA")
```

Get two sequences.

```{r}
x <- as.character(or[[1]])
y <- as.character(or[[2]])
x # case doesn't matter
```

Get the substitution matrix.

```{r}
mtx <- get_substitute_mtx("PAM250")
```

Calculate weighted distance between the two sequences.

```{r}
weights <- rep(1, length(x)) # ideally should reflect your belief on the importance of each residue, all 1 here
two_seq_score(x, y, mtx, weights)
```

The hope is to be able to calculate pairwise distances of larger number of sequences. Pairwise distances for all mouse olfactory receptors:

```{r}
lst <- purrr::map(or, ~as.character(.x))
names(lst) <- purrr::map(or, ~attr(.x, "Annot"))
dist_mtx <- pairwise_score(lst, mtx, weights)
```

This will actually lead to some clustering of receptors (e.g. compare with Figure 2 here http://stke.sciencemag.org/content/2/60/ra9)

```{r}
dist_pca <- prcomp(dist_mtx, scale. = TRUE)
ggplot2::ggplot(as.data.frame(dist_pca$x), ggplot2::aes(PC1, PC2)) +
  ggplot2::geom_point(alpha=0.3)
```

We can also try ~10k sequences, the time should still be reasonable. There's also a `parallel_pairwise_score` to be used when there are multiple cores available. On my 2015 Macbook Pro they took ~ 1 minute and ~ 30 seconds respectively.

```{r}
long_lst <- rep(lst, 10)
names(long_lst) <- 1:length(long_lst)
long_weights <- rep(1, length(long_lst))
system.time(dist_mtx <- pairwise_score(long_lst, mtx, weights))
system.time(dist_mtx_p <- parallel_pairwise_score(long_lst, mtx, weights))
identical(dist_mtx, dist_mtx_p)
```

You can further benchmark if interested. I'm not running this part.

```{r, eval=FALSE}
microbenchmark::microbenchmark(dist_mtx <- pairwise_score(lst, mtx, weights))
microbenchmark::microbenchmark(dist_mtx_p <- parallel_pairwise_score(lst, mtx, weights))
```
